---
title: "lab9"
author: "Reed Magleby"
date: "November 7, 2018"
output: pdf_document
---

1. Consider a function that counts number of odd numbers in a given vector

```{r}
odd_count = function(x) {
odd_num = 0
for (i in 1:length(x)) {
if (x[i] %% 2 == 1) odd_num = odd_num + 1
}
return(odd_num)
}

odd_count(c(1,2,3,4,5,6,7,8,9))
```


Write a vectorized form of this function and compare their performance using microbenchmark function
from the package microbenchmark.

```{r}
odd_vect = function(x) {
  odd_num = sum(x %% 2 == 1)
  return(odd_num)
}

odd_vect(c(1,2,3,4,5,6,7,8,9))
```

```{r}
test <- seq(0, 1000, by=1)

library(microbenchmark)
microbenchmark(odd_vect(test),odd_count(test))

```


2. Create a matrix and calculate sum of each column:
. using loop
. using apply() function
. using colSums() function
Compare performance of the functions using microbenchmark() function

```{r}
test2 <- matrix(seq(1, 100, by = 1), byrow = TRUE, nrow = 10)

colSums(test2)

for_sum <- function(x){
  sum_matrix <- matrix()
  for (i in 1:ncol(x)){ 
    sum_matrix[i] <- sum(x[,i])
  }
  return(sum_matrix)
}

apply_sum <- function(x){
  apply(x,2,sum)
}

microbenchmark(colSums(test2), for_sum(test2), apply_sum(test2))

```


3. Create a random vector of positive integers of length 1000000 elements. Use profvis() function to
profile the odd_count() function defined in the problem 1 with the vector as an input.

```{r}
library(profvis)
test3 <- sample.int(10, 1000000, replace = TRUE)

profvis({
  odd_count <- function(x) {
    odd_num <- 0
    for (i in 1:length(x)) {
      if (x[i] %% 2 == 1) odd_num <- odd_num + 1
    }
    return(odd_num)
  }
  odd_count(test3)
})

```
2. Modify the sorting function (sort_vec) from "Assignment 8" (problem 3) so that it should take an
additional argument ascending which causes sorting in increasing order when 'ascending = TRUE'. In
other words,
. sort_vect(c(3, 1, 2), ascending = TRUE) = (1, 2, 3)
. sort_vect(c(3, 1, 2), ascending = FALSE) = (3, 2, 1)


```{r}
sort_vec <- function(x, ascending = TRUE) {
  if (length(x) < 2) return(x)
  for (last in length(x):2) {
    for (first in 1:(last - 1)) {
      if (x[first] > x[first + 1]) {
        temp <- x[first]
        x[first] <- x[first + 1]
        x[first + 1] <- temp
      }
    }
  }
  if (ascending == TRUE) return(x) else return(rev(x))
}

sort_vec(c(3,1,2,4), FALSE)
```

3. Consider a simple random walk with starting point 0 and a step -1 or 1. Below is the code with
dynamically allocated memory. Write your code with preallocated memory and compare time for both
versions using system.time() function (use N = 1000, 10000 and 1000000).

```{r}
for (j in (c(1000, 10000, 1000000))) {
  data_series <- 0
  ans <- system.time({
    for (i in 2:j) {
      data_series[i] <- data_series[i - 1] + sample(c(-1, 1), 1)
    }
  })
  print(ans)
}
```


```{r}

for (j in (c(1000, 10000, 1000000))) {
  data_series <- vector(length = j)
  ans <- system.time({
    for (i in 2:j) {
      data_series[i] <- data_series[i - 1] + sample(c(-1, 1), 1)
    }
  })
  print(ans)
}
```



